\textbf{NBA Party}

### 2017-07-08 cardinal_analytx_presentation/nightlife_and_sports.tex
Changed where 'prev research' slides placed. Updated tables and formatting.

    Conditions on (almost) every conceivable element of the game:
    \begin{itemize}     \item strengths and weaknesses against particular opponents,
    \item momentum,
    \item injuries, and
    \item even possibly a weather forecast.     \end{itemize}

@@

\begin{frame}   \frametitle{Partying as a latent variable}
  \begin{block}{We can't observe when players go for a night on the town.}
    Instead, we develop two approaches to proxy for partying.

@@

\begin{frame}   
  \frametitle{Continuous measure of party}
  \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location} 
\end{frame}

\begin{frame}   \frametitle{Results: NBA}
  \vspace{-2pt}
  \centering
  \tiny{ % m0, m1
\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.923 & $-$0.875 \\    & (0.695) & (0.695) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf. Crit. & 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular} 
}   
\end{frame}

\begin{frame}   \frametitle{Interpreting our results}
  Let's consider the simplest model, where $Y_i | X \sim \textrm{Bern}(p)$, i.e.
  $\Pr(Y_i = y | X) = p^{y} (1-p)^{1-y}$. Using a generalized linear model,
  we fit a linear predictor to a transformation of the expected value of our
  variable of interest. I.e.
  \[
    \textrm{logit} \left( \mathbb E[Y_i | X]\right) = \textrm{logit} \left(p\right) = \ln \left( \frac{p}{1-p}\right) = X \beta.
  \]
  So, we interpret our regression coefficient as follows: visiting LA or NY within
  24 hours of another game corresponds to being $e^{\beta_1} \approx e^{-0.55} \approx \frac{3}{5}$
  as likely to meet the spread, all other factors held constant. 
\end{frame}

@@

\begin{frame}   \frametitle{Defense in the NBA}
  \begin{adjustwidth}{-.75cm}{}
  \tiny{  % mpa, tpa, tps
\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std. Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}      }
\end{adjustwidth} 
\end{frame}

@@

\begin{frame}   \frametitle{Conclusions}
  \begin{itemize}     \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY, and in general we see that performance decreases as a function of (our proxy for) nightlife.
    \item The decrease in likelihood can (in part) be explained by a drop in defensive performance.
    \item With a causal model established, we realize a profitable betting scheme.   \end{itemize} \end{frame}

>>>>

    Conditions on (almost) every conceivable element of the game:
    \begin{itemize}     \item strengths and weaknesses against particular opponents,
    \item momentum,
    \item injuries, and
    \item even possibly a weather forecast (for outdoor sports).     \end{itemize}

@@

\begin{frame}   \frametitle{Recall: partying is a latent variable}
  \begin{block}{We can't observe when players go for a night on the town.}
    Instead, we develop two approaches to proxy for partying.

@@

\begin{frame}   
  \frametitle{Continuous measure of party}
  \centering \includegraphics[scale=0.375]{../writing/Party_by_Team_Location} 
\end{frame}

\begin{frame}   \frametitle{Results: NBA}
  \vspace{-2pt}
  \centering
  \tiny{ % m0, m1
\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf. Crit. & 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }   
\end{frame}

\begin{frame}   \frametitle{Interpreting our results}
  Let's consider the simplest model, where $Y_i | X \sim \textrm{Bern}(p)$, i.e.
  $\Pr(Y_i = y | X) = p^{y} (1-p)^{1-y}$. Using a generalized linear model,
  we fit a linear predictor to a transformation of the expected value of our
  variable of interest. I.e.
  \[
    \textrm{logit} \left( \mathbb E[Y_i | X_i]\right) = \textrm{logit} \left(p\right) = \ln \left( \frac{p}{1-p}\right) = X_i^T \beta.
  \]
  So, we interpret our regression coefficient as follows: visiting LA or NY within
  24 hours of another game corresponds to being $e^{\beta_1} \approx e^{-0.55} \approx \frac{3}{5}$
  as likely to meet the spread, all other factors held constant. 
\end{frame}

@@

\begin{frame}   \frametitle{Defense in the NBA}
  \begin{adjustwidth}{-.75cm}{}
  \tiny{  % mpa, tpa, tps
\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std. Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
\end{adjustwidth} 
\end{frame}

@@

\begin{frame}   \frametitle{Conclusions}
  \begin{itemize}     \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY.
    \item In general we see that performance decreases as a function of (our proxy for) nightlife.     \item The decrease in likelihood can (in part) be explained by a drop in defensive performance in the NBA.
    \item With a causal model established, we realize a profitable betting scheme in MLB.   \end{itemize} \end{frame}
